---
title: "PML project"
author: "Parikshit Sanyal"
date: "4 September 2018"
output: html_document
---

---
title: "Prediction of exercise type from physical measurements"
author: "Parikshit Sanyal"
date: "31 August 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(RANN)
```

## Loading the data

We load the training dataset in the variable 'tr'. This is the dataset we are later going to split into 'training' and 'testing'. The validation data 'val' (from pml-testing.csv) is going to be untouched till the last phase.

```{r loaddata}
tr <- read.csv('pml-training.csv')
val <- read.csv('pml-testing.csv')
```

We take a look at how big is the 'tr' dataset.

```{r trdata}
dim(tr)
```

... which shows it is compiled of 19622 records of 160 variables. We take a look at the 'classe' variable which we have to predict

```{r classe}
summary(tr$classe)
```

It is a factor of 5 levels A, B, C, D and E. We now try to visualise relations between the variables. We check for NA values in the set

```{r}
sum(is.na(tr))
```

We need to impute these NA values by K nearest eighbours algorithm

```{r impute, eval=FALSE}
preProcValues <- preProcess(tr, method = c("knnImpute","center","scale"))
tr_processed <- predict(preProcValues, tr)
sum(is.na(tr_processed))
```

We use readings from accelometers in the belt , forearm, arm and dumbbell; we segregate these variables

```{r accelnames}
tr_processed= tr
names(tr_processed)[grep("accel",names(tr_processed))]
ta <- names(tr_processed)[grep("^total_accel",names(tr_processed))]
```
We now plot total accels against each other

```{r plottingpredictors}
featurePlot(x = tr_processed[,ta],y=tr$classe,plot='pairs')

```
Since the total_accels in each recorder do not show any correlation with each other, we can assume them to be independent variables, and use them as predictors.

```{r partition}
set.seed(107)
inTrain <- createDataPartition(
  y = tr_processed$classe,
  p = .75,
  list = FALSE
)
training <- tr_processed[ inTrain,]
testing  <- tr_processed[-inTrain,]
nrow(training)
nrow(testing)
```

We now train a KNN based model from the dataset and predict values from testing set

```{r trn, eval=FALSE}
val_acc <- val[,ta]
model_knn <- knn3(classe ~ .,data=training,k=3)
pt_k <- predict(model_knn, testing, type='class')
summary(testing$classe == pt_k)
```

We then apply the model to the valuation set
```{r val,eval=FALSE}
pv_k <- predict(model_knn,val_acc,type='class')
pv_k
```

Which outputs E B E B E B B B E E E B E E B E E E B B for the 20 entries in val dataset.